---
title: "CUNY SPSS Data Science Math Homework 1"
author: "Mike Silva"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(Deriv)
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

Profit is the difference between revenue and cost.  Revenue is the quantity sold (q) multiplied by the price(p) and the costs is given at $10 times the quantity sold, thus
$$P(q,p) = R(q,p) - C(q) = qp - 10q$$
The quantity sold is given as a function of the price
$$q(p) = 80-2p$$
So rewriting the equation for profit as a function of the price
$$P(p) = R(p) - C(p)$$
$$ (80-2p)p - 10(80-2p)$$
$$ (80p-2p^2) - (800-20p)$$
$$-2p^2 + 100p - 800 $$

******

### Question 2

Given $f(x)=8x^3+7x^2-5$:

#### Step 1

$$f(3) = 8(3)^3+7(3)^2-5 = (8\cdot27)+(7\cdot9)-5 = 216+63-5 = 274$$

#### Step 2

$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8\cdot-8)+(7\cdot4)-5 = -64+28-5 = -41$$

#### Step 3

$$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$
Verifying my math with R:
```{r}
f2 <- function(x){8*x**3+7*x**2-5}
f2(3)
f2(-2)
```

******

### Question 3

#### Step 1

$$\lim_{x\to1^-} f(x) = 2$$

#### Step 2

$$\lim_{x\to1^+} f(x) = -5$$

#### Step 3

$$\lim_{x\to1} f(x) = \mbox{ Does not exist} $$

******

### Question 4

Find the derivative of $f(x)=-2x^3$

$$f'(x)=-6x^2$$
Verifying my math in R:
```{r}
f4 <- function(x){-2*x**3}
Deriv(f4)
```

******

### Question 5

$$f(x) = \frac{-8}{x^2}$$
$$f'(x) = \frac{16}{x^3}$$
Verify with R:
```{r}
f5 <-function(x){-8/(x**2)}
Deriv(f5)
```

******

### Question 6
$$g(x) 5\sqrt[3]{x}$$

$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$
```{r}
f6 <- function(x){5*x**(1/3)}
Deriv(f6)

5/3

2/3
```

******

### Question 7

$$y=-2x^\frac{9}{8}$$

$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$
```{r}
f7 <- function(x){-2*x**(9/8)}
Deriv(f7)

-9/4
1/8
```

******

### Question 8

$$\frac{35-40}{4-0}=\frac{-5}{4} = -1\frac{1}{4}$$

******

### Question 9

$$C(x)=630+2.4x$$
$$\mbox{Avg Cost} = \frac{\mbox{Cost}}{\mbox{Unit}}=\frac{C(x)}{x}=\frac{630+2.4x}{x}$$

******

### Question 10

$$f(x)=(-2x^{-2} +1)(-5x + 9)$$

Applying the quotient rule where $f(x)=-2x^{-2}+1$ and $g(x)=(-5x+9)$

$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$
Checking my math with R:
```{r}
f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
Deriv(f10)
# Evaluating at 10
f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
f10_prime(10)
f10_prime <- Deriv(f10)
f10_prime(10)
```

******

### Question 11

$$f(x)=\frac{5x^\frac{1}{2}+7}{x^3 + 1}$$

Applying the quotient rule where $f(x)=5\sqrt{x}+7$ and $g(x)=x^3+1$ so $f'(x)=\frac{5}{2\sqrt{x}}$ and $g'(x)=3x^2$

$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5\sqrt{x}+7))}{(x^3+1)^2}$$
$$f'(x) = \frac{\frac{5x^3+5}{2\sqrt{x}}-3x^2 (5\sqrt{x}+7)}{(x^3+1)^2}$$

Simplified to 
$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$
Again checking with R:
```{r}
f11 <- function(x){(5*x**(1/2)+7)/(x**3+1)}
Deriv(f11)
# Evaluate at 5
fprime11 <- function(x){-1*(42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*((x**3+1)**2))}
fprime11(5)
fprime11 <- Deriv(f11)
fprime11(5)
```

******

### Question 12

$$f(x)=(3x^{-3}-8x+6)^\frac{4}{3}$$

$$f'(x) = \frac{4}{3}(-\frac{9}{x^4}-8) \Big(\frac{3}{x^3}-8x+6\Big)^\frac{1}{3}$$

Checking my math in R:
```{r}
f12 <- function(x){(3*x**-3-(8*x)+6)**(4/3)}
Deriv(f12)
# Evaluate at 1
f12_prime <- function(x){(4/3)*((-9/x**4)-8)*(((3/x**3)-8*x+6)**(1/3))}
f12_prime(1)
f12_prime <- Deriv(f12)
f12_prime(1)
```

******

### Question 13

$$f(t) =\frac{550t^2}{\sqrt{t^2+15}}$$

After factoring out the 550 constant and applying the quotient rule where $f(x)=t^2$ and $g(x)=\sqrt{t^2+15}$

$$550\cdot f'(x) = \frac{(2t)(\sqrt{t^2+15})-(t^2)\Big(\frac{t}{\sqrt{t^2+15}}\Big)}{t^2+15} = \frac{2t\sqrt{t^2+15} -\frac{t^3}{\sqrt{t^2+15}}}{t^2+15}$$

Multiplying by 550 and simplifying
$$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$
```{r}
f13 <- function(t){(550*t**2)/((t**2+15)**(1/2))}
Deriv(f13)
f13_prime <- function(t){(550*t*(t**2+30))/((t**2+15)**(3/2))}
f13_prime(3)
f13_prime <- Deriv(f13)
f13_prime(3)

```

The answer is 547

******

### Question 14

Given $N(t) = 1000(6+0.1t)^\frac{1}{2}$
$$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$