---
title: "DATA 605 Week 7 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

## Question 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

### Part A

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a geometric.**

The probability the machine will fail within ten years is 1.  The probability that the machine will fail within k years is found using:

$$P(x=k) = (1-p)^{k-1} \times p$$
where p is the probability of the event happening that year.  We are given that p is 1/10.  The probability it will fail in the first eight years is displayed below

```{r echo=FALSE}
library(ggplot2)
n_failures_before_success <- 8
probability_of_success <- 1/10

p_ns <- c()
for (n in 1:n_failures_before_success){
  p_n <- probability_of_success * (1 - probability_of_success)^(n-1)
  p_ns <- c(p_ns, p_n)
}

ggplot(data.frame(year = 1:n_failures_before_success, p = p_ns), aes(year,p)) + 
  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
  geom_text(aes(label = round(p,3)), vjust = 1.6, color = "white", size = 3.5) +
  theme_minimal() +
  ggtitle("Probability of Failure in the First 8 Years")
```

The probability of failure after the eight years will be one minus the sum of all the probabilities above (the area in blue).  Otherwise stated:

$$P(\text{failure after 8 years}) = 1 - \sum_{k=1}^{8} \frac{1}{10}\times\Big(\frac{9}{10}\Big)^{k-1} = 0.4304672$$

```{r}
n_failures_before_success <- 8
probability_of_success <- 1/10

p <- 0
for (n in 1:n_failures_before_success){
  p <- p + probability_of_success * (1 - probability_of_success) ^ (n - 1)
}

p_geometric <- 1 - p

message(paste("Probability that the machine will fail after 8 years:", p_geometric))
```

The [expected value](https://math.stackexchange.com/questions/2368304/how-do-you-calculate-the-expected-value-of-geometric-distribution-without-diffre) of a geometric distribution is 

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

The [standard deviation](https://www.youtube.com/watch?v=kDFiNt8nxSo) is:

$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$

```{r}
variance <- (1 - probability_of_success) / (probability_of_success ^ 2)

standard_deviation <- sqrt(variance)

message(paste("Standard deviation:", standard_deviation))
```

### Part B

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as an exponential.** 

The probability the machine fails after 8 years is 

$$P(\text{failure after 8 years}) = e^\frac{-8}{10} = e^{-0.8}=0.449329$$

```{r}
p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value)

message(paste("Probability that the machine will fail after 8 years:", p_exponential))
```

The expected value is found using the same formula as above as demonstrated in [this video](https://www.youtube.com/watch?v=GJ2klfD0Q3g). Therefore

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

The [standard deviation](https://www.youtube.com/watch?v=GJ2klfD0Q3g) is:

$\text{Standard Deviation} = \frac{1}{P(success)^2} = \frac{1}{0.1^2} = \frac{1}{0.01} = 100$

```{r}
standard_deviation <- 1 / (probability_of_success ^ 2)

message(paste("Standard deviation:", standard_deviation))
```

### Part C

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial.**



## Part D

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a Poisson**


## Summary

|        | Probability of Failure After 8 Years | Expected Value | Standard Deviation |
| Part A | 0.4304672                            | 10             | 9.48683298050514   |
| Part B | 0.449329                             | 10             | 100                |
| Part C |
| Part D |