---
title: "DATA 605 Week 8 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 11 on Pg. 303

**A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?**

The exponential formula is:

$$N(t) = N_0 e ^ {-\lambda t} $$

lambda would be 1 failure every 1000 hours.  Otherwise stated: 

$$\lambda = \frac{1}{1000}$$

Now we need to plug in the variables and solve for t 

$$100 e ^ {-\frac{1}{1000}t} = 99 \rightarrow e ^ {-\frac{1}{1000}t} = \frac{99}{100} \rightarrow -\frac{1}{1000}t = ln\bigg(\frac{99}{100}\bigg) \rightarrow t = ln\bigg(\frac{99}{100}\bigg)\times-1000 \approx 10.05034$$

**Answer: 10.05034 hours**

## Question 14 on Pg. 303

**Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density $fz(z)=\frac{1}{2}\lambda e ^{-\lambda \mid z \mid}$**

*Note: In order to keep this straigth in my head I solved Z = X - Y instead of Z = X1 - X2.  This notation will be used in this solution.*

The convolution formula of Z = X + Y is

$$fx(z) = \int_{-\infty}^{\infty}fx(x)fy(z-x)dx$$

But to do Z = X - Y we need to do Z = X + (-Y).  Pluging it into the convolution formula:

$$fz(z) = \int_{-\infty}^{\infty}fx(x)f{-y}(z-x)dx$$
Since $f{-y}(z-x) = fy(x-z)$ we can rewrite the convolution as

$$fz(z) = \int_{-\infty}^{\infty}fx(x)fy(x-z)dx$$

Looking at the case where z < 0

$$ \int_{0}^{\infty}\lambda e ^{-\lambda x} \lambda e ^ {-\lambda(x-z)}dx = \lambda e ^ {\lambda z}\int_{0}^{\infty}\lambda e ^{-2\lambda x} dx = \lambda e ^{\lambda z}(-\frac{1}{2}e ^ {-\lambda x}|_0^{\infty}) = \frac{1}{2}\lambda e^{\lambda z}$$

Now looking at the case where z is greater than or equal to zero.  Since the two variables are independent Z = X - Y is the same as -Z = -X + Y.  So the distribution is symetric arround zero thus fz(z) = fz(-z). Sp when z >= 0:

$$fz(-z) = \frac{1}{2}\lambda e^{-\lambda z}$$

This implies that $fz(z)=\frac{1}{2}\lambda e ^{-\lambda \mid z \mid}$

## Question 1 Pg. 320-321

**Let X be a continuous random variable with mean µ = 10 and variance $\sigma^2$ = 100/3. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.**

```{r}
mu <- 10
sigma <- sqrt(100/3)
```

### (a) P(|X − 10| ≥ 2).

```{r}
k <- 2 / sigma
1 / k^2
```

### (b) P(|X − 10| ≥ 5).

```{r}
k <- 5 / sigma
1 / k^2
```

### (c) P(|X − 10| ≥ 9).

```{r}
k <- 9 / sigma
1 / k^2
```

### (d) P(|X − 10| ≥ 20).

```{r}
k <- 20 / sigma
1 / k^2
```