---
title: "DATA 605 Assignment 2"
author: "by Michael Silva"
output: html_document
---

## Problem Set 1

**(1) Show that $A^TA \neq AA^T$ in general.**

\[ A=\begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix}
\]

\[ A^T=\begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix}
\]

\[ A^TA=
  \begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix} \times
  \begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix} = \begin{bmatrix}
  a \times a + c \times c & c \times a + c \times c \\
  b \times b + d \times d & b \times d + d \times d
  \end{bmatrix} = \begin{bmatrix}
  a^2 + c^2 & ac + c^2 \\
  b^2 + d^2 & bd + d^2
  \end{bmatrix} 
\]

\[ AA^T=\begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix} \times
  \begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix} = \begin{bmatrix}
  a \times a + b \times b & a \times c + b \times d \\
  c \times a + d \times d & c \times c + d \times d
  \end{bmatrix} = \begin{bmatrix}
  a^2 + b^2 & ac + bd \\
  ac + d^2 & c^2 + d^2
  \end{bmatrix} 
\]

\[
\begin{bmatrix}
  a^2 + c^2 & ac + c^2 \\
  b^2 + d^2 & bd + d^2
  \end{bmatrix} 
  \neq
  \begin{bmatrix}
  a^2 + b^2 & ac + bd \\
  ac + d^2 & c^2 + d^2
  \end{bmatrix} 
\]

**(2) For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions could this be true?**

Any matrix where all elements are the same value.  For example the zero matrix:

\[ A=\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix}
\]

\[ A^T=\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix}
\]

\[ A^TA=
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} \times
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} = \begin{bmatrix}
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0 \\
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0
  \end{bmatrix} = \begin{bmatrix}
  0 + 0 & 0 + 0 \\
  0 + 0 & 0 + 0
  \end{bmatrix} 
  = \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

\[ AA^T=
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} \times
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} = \begin{bmatrix}
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0 \\
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0
  \end{bmatrix} = \begin{bmatrix}
  0 + 0 & 0 + 0 \\
  0 + 0 & 0 + 0
  \end{bmatrix} 
  = \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

\[
\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
  =
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

## Problem Set 2

**Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.**

**Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention.**

**You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems. **

```{r}
factorize <- function(A){
  n <- nrow(A)
  U <- A
  L <- diag(n)

  for (j in c(1:n)){
    for(i in c(2:n)){
      if(i > j){
        the_row <- U[j,]
        multiplier <- U[i, j] / the_row[j]
        U[i,] <- U[i,] - (multiplier * the_row)
        L[i,j] <- multiplier
      }
    }
  }
  
  return(list(L=L, U=U))
}
```

I will test this on the example worked through on [this video](https://www.youtube.com/watch?v=UlWcofkUDDU&feature=youtu.be)

$A = LU$ or

\[ \begin{bmatrix}
  2 & 4 & -4 \\
  1 & -4 & 3\\
  -6 & -9 & 5
  \end{bmatrix} = 
  \begin{bmatrix}
  1 & 0 & 0 \\
  \frac{1}{2} & 1 & 0\\
  -3 & -\frac{1}{2} & 1
  \end{bmatrix} \times
  \begin{bmatrix}
  2 & 4 & -4 \\
  0 & -6 & 5\\
  0 & 0 & -\frac{9}{2}
  \end{bmatrix}
\]

```{r}
A <- matrix(c(2, 1, -6, 4, -4, -9, -4, 3, 5), 3, 3)

factors <- factorize(A)
factors$L
factors$U
```