---
title: "DATA 605 Week 15"
author: "Mike Silva"
date: "May 6, 2019"
output: html_document
---

## Question 1

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

Given the set of (X, Y) points we need to find the best fit line $\hat{Y}= a + bX$ such that the sum of square errors in Y ($\sum{(Y-\hat{Y})^2}$) is minimized.

$$SSE=\sum_{i=1}^{n}{(Y_i-\hat{Y})^2} = \sum_{i=1}^{n}{(Y_i-a - bX_i)^2}$$

SSE will be minimized when $\frac{d{SSE}}{da} = 0$ and $\frac{d{SSE}}{db} = 0$.


$$\frac{d{SSE}}{da} = \sum_{i=1}^{n}{-2(Y_i-a - bX_i)}=2\bigg(na+b\sum_{i=1}^{n}{X_i}-\sum_{i=1}^{n}{Y_i}\bigg)=0$$

Dividing by 2 and solving for a

$$a = \bar{Y}-b\bar{X}$$

$$\frac{d{SSE}}{db} = \sum_{i=1}^{n}{-2X_i(Y_i-a - bX_i)}=\sum_{i=1}^{n}{-2(X_iY_i-aX_i-bX_i^2)}=0$$

Substituting in a gives

$$\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y}-bX_i\bar{X}-bX_i^2)}=0 \rightarrow \sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}-b\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}=0$$

Solving for b gives

$$b=\frac{\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}}{\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}}$$

```{r}
X <- c(5.6, 6.3, 7, 7.7, 8.4)
Y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

X_bar <- mean(X)
Y_bar <- mean(Y)

b <- sum(X * Y - X * Y_bar) / sum(X ^ 2 - X * X_bar)

a <- Y_bar - b * X_bar
```

The least squares line is $\hat{Y}$ = `r a` + `r b` $X$

Double checking my work

```{r}
lm(Y ~ X)
```

## Question 2

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.

$$f(x,y)=24x-6xy^2-8y^3$$

First we find the first and second partial derivatives

| First Derivative    | Second Derivative     |
|---------------------|-----------------------|
|$f_x = 24 - 6y^2$    | $f_{xx} = 0$          |
|$f_y = -12xy -24y^2$ | $f_{yy} = -12x - 48y$ |

and $f_{xy} = 12y$.  Now we find the critical points

$$f_x(x,y) = 0 \rightarrow 24 - 6y^2=0 \rightarrow -6y^2 = -24 \rightarrow y^2 = 4 \rightarrow y = \pm 2 $$

$$f_y(x,y) = 0 \rightarrow  -12xy - 24y^2 = 0 \rightarrow-12xy = 24y^2 \rightarrow  x = -2y$$

When y = 2 x = -4 and when y = -2 x = 4.  In order to find the z we plug these values into the function.

```{r}
fxy <- function(x, y){
  24*x-6*x*y^2-8*y^3
}

fxy(4, -2)
fxy(-4, 2)
```

In order to figure out what type of point it is, we need to plug it into the second derivative test.

```{r}
second_derivative_test <- function(x,y){
  #fxx * fyy - fxy^2
  (0 * (-12*x - 48*y))-((12*y)^2)
}

second_derivative_test(4, -2)
second_derivative_test(-4, 2)
```

| Point (x, y, z) | Second Derivative Test | Conclusion   |
|-----------------|------------------------|--------------|
|(4, -2, 64)      | Negative               | Saddle Point |
|(-4, 2, -64)     | Negative               | Saddle Point |

## Question 3

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x -23y units of the "name" brand.

**Step 1.** Find the revenue function R (x, y).

$R(x,y) = x(81 - 21x + 17y) + y(40 + 11x - 23y)$

**Step 2.** What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}
revenue <- function(x,y){
  x * (81 - 21 * x + 17 * y) + y * (40 + 11 * x -23 * y)
}

revenue(2.30, 4.10)
```

R(x,y) = $116.62

## Question 4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x,y)=\frac{1}{6}x^2+\frac{1}{6}y^2+7x+25y+700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

Let's express the cost equation as a function of x.  Given $x + y = 96$, $y = 96 - x$.

$$C(x)=\frac{1}{6}x^2+\frac{1}{6}(96-x)^2+7x+25(96-x)+700$$

$$C(x)=\frac{x^2}{6}+\frac{1}{6}(96-x)^2-18x+3100$$

$$C(x)=\frac{x^2}{3}-50x+4636$$

Now minimize it

$$C'(x) =  \frac{2x}{3}-50 \rightarrow 0 = \frac{2x}{3}-50 \rightarrow\frac{2x}{3} = 50 \rightarrow 2x = 150 \rightarrow x = 75$$

$$y = 96 - (75) \rightarrow y = 21$$

| Location    | Units |
|-------------|-------|
| Los Angeles | 75    |
| Denver      | 21    |

## Question 5

Evaluate the double integral on the given region.

$\int{\int{(e^{8x+3y})dA}}$; $R: 2 \leq x \leq 4$ and $2 \leq y \leq 4$

Write your answer in exact form without decimals.

$$A = \int_2^4{\int_2^4{e^{8x+3y}}}$$

$$\int{e^{8x+3y}dx} = \frac{1}{8}e^{8x+3y} + C |_2^4 = \frac{1}{8}e^{8(4)+3y} - \frac{1}{8}e^{8(2)+3y} = \frac{1}{8}e^{32+3y} - \frac{1}{8}e^{16+3y} = \frac{e^{16}-1}{8}e^{3y+16}$$

$$\int{\frac{e^{16}-1}{8}e^{3y+16}dy}=\frac{e^{16}-1}{24}e^{3y+16} + C|_2^4 = \frac{e^{44}-e^{28}}{24} - \frac{e^{38}-e^{22}}{24}$$
