---
title: "CUNY SPS DATA-607 Week 1 Assignment"
subtitle: "Loading Data into a Data Frame"
author: "Mike Silva"
date: "August 27, 2018"
output: html_document
---

## Description

The assignment is to load the Mushroom Dataset from the [UCI repository](https://archive.ics.uci.edu/ml/datasets/Mushroom) into a data frame, rename the columns into something meaningful, replace the abbreviations used in the data, and subset the columns of the data frame.  It must include the column that indicated if the mushroom is edible or poisonous and three or four other columns.

## Step 1: Acquire the Data

The first step will be to download the data to the local environment:

```{r acquire-data}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'agaricus-lepiota.data')
```

## Step 2: Load the Data into Data Frame

Now that we have the dataset locally we will create a data frame:

```{r load-data}
df <- read.table('agaricus-lepiota.data', sep=',')
```

According to the documentation we should have 8,124 rows and 22 columns of data.

```{r data-size}
dim(df)
```

## Step 3: Rename the columns

Our data checks out, but the names are not very useful.

```{r bad-names}
names(df)
```

I will rename the columns by using the dplyr package:

```{r rename-columns, warning=FALSE}
library(dplyr)

df <- df %>%
  r
```

