---
title: "DATA 621 Homework #2"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/classification-output-data.csv")
actual <- "class"
predicted <- "scored.class"
probability <- "scored.probability"
# Load our functions
source("hw2.R")
```

## Caret's Output

```{r}
confusionMatrix(factor(df[[predicted]]), factor(df[[actual]]))
```

## Our Function's Output 

**Confusion Matrix**

```{r}
confusion_matrix(df, actual, predicted) %>%
  kable() %>%
  kable_styling()
```

**Accuracy**: `r accuracy(df, actual, predicted)`

**Classification Error Rate**: `r classification_error_rate(df, actual, predicted)`

**Precision**: `r precision(df, actual, predicted)`

**Sensitivity (Recall)**: `r sensitivity(df, actual, predicted)`

**Specificity**: `r specificity(df, actual, predicted)`

**F1 Score**: `r f1_score(df, actual, predicted)`

### F1 Score Will Always be Between 0 and 1
Before we move on, let’s consider a question that was asked: What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0<a<1 and 0<b<1 then ab<a).

  
**Proving the range of the F1 Score must be between 0 and 1**
  
**Let's expand the F1 equation using the definitions of precision and sensitivity.**


$$\textrm{F1 Score}= \frac{2 \times\mathrm{Precision}\times\mathrm{Sensitivity}}{\mathrm{Precision}+\mathrm{Sensitivity}}$$ 
**The numerator is equal to**
$$=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)}$$ 
**The denominator is equal to**
$$= \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
**Combine the Numerator and the denominator:**
$$F1=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)} \times \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
**Multiply the denominator by (TP + FP)(TP + FN):**

$$denominator = \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
$$F1=\frac{2(TP)^2}{(TP+FP)(TP+FN)}\times \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
**Cancel the denominator of the first term with the numerator of the second term:**

$$F1 =\frac{2(TP)^2}{TP(TP+FN)+TP(TP+FP)}$$
**Simplify the Denominator:**
$$F1 =\frac{2(TP)^2}{TP^2+FNTP+TP^2+TPFP}$$  
$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
$$F1 =\frac{2(TP)^2}{TP(2TP +FN +FP)}$$ 
**Cancel the TP terms and get final answer**
$$F1=\frac{2TP}{2TP+FN+FP}$$
<br>  
In the equation above, $TP,FN,FP \in \mathbb{N}$ where $\mathbb{N} = \{0,1,2,3...\}$  
 
$2TP \le2TP+FN+FP$ this equation shows that the numerator is at most equal to the denominator (i.e $F1\le1$).  
  
There would be a Maximum value of $1$ when $FN=FP=0$ and $TP > 0$  

The fraction $=\frac{2TP}{2TP+FN+FP}$ has a minimum value of zero when the numerator is minimized (i.e. $TP=0$).  
  
We have shown that $0\le F1\le1$
<br>  

## pROC Output

```{r}
par(pty = "s")
roc(df[[actual]], df[[probability]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

## Our ROC Curve

```{r}
our_roc <- roc_curve(df, actual, probability)

our_roc$plot
```

```{r}
our_roc$auc_vector
```

## Source Code

```{r source, code=readLines("hw2.R")}
```