---
title: "DATA 624 Homework 1"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---


```{r setup, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
```

## 2.1 

Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.

a. Use `autoplot()` to plot each of these in separate plots.

b. What is the frequency of each series? Hint: apply the `frequency()` function.

c. Use `which.max()` to spot the outlier in the gold series. Which observation was it?

## 2.2 

Download the file `tute1.csv` from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a. You can read the data into R with the following script:

```{r}
download.file("http://otexts.com/fpp2/extrafiles/tute1.csv", "tute1.csv")
tute1 <- read.csv("tute1.csv", header=TRUE)
View(tute1)
```

b. Convert the data to time series

```{r}
mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```

(The [,-1] removes the first column which contains the quarters as we don’t need them now.)

c. Construct time series plots of each of the three series

```{r}
autoplot(mytimeseries, facets=TRUE)
```

Check what happens when you don’t include facets=TRUE.

```{r}
autoplot(mytimeseries)
```

## 2.3

Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

a. You can read the data into R with the following script:

```{r}
download.file("https://otexts.com/fpp2/extrafiles/retail.xlsx", "retail.xlsx")
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

The second argument (skip=1) is required because the Excel sheet has two header rows.

b. Select one of the time series as follows (but replace the column name with your own chosen column):

```{r}
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
```

c. Explore your chosen retail time series using the following functions:

`autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()`

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

## 2.6

Use the following graphics functions: `autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

* Can you spot any seasonality, cyclicity and trend?

* What do you learn about the series?
