---
title: "DATA 624 Homework 1"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---


```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
library(gridExtra)
```

## Question 2.1 

Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.

**`gold` is the daily morning gold prices in US dollars from 1 Jan 1985 to 31 March 1989.  `woolyrnq` is the quarterly production of woolen yarn in Australia in tons from March 1965 to September 1994.  `gas` is monthly gas production in Australia from 1956-1995.**

a. Use `autoplot()` to plot each of these in separate plots.

```{r}
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

b. What is the frequency of each series? Hint: apply the `frequency()` function.

```{r}
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

c. Use `which.max()` to spot the outlier in the gold series. Which observation was it?

```{r}
goldoutlier <- which.max(gold)
```

**It is the `r goldoutlier` observation.  The price of gold was `r gold[goldoutlier]`**

## Queston 2.2 

Download the file `tute1.csv` from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a. You can read the data into R with the following script:

```{r}
if(!file.exists("tute1.csv")){
  download.file("http://otexts.com/fpp2/extrafiles/tute1.csv", "tute1.csv")
}
tute1 <- read.csv("tute1.csv", header=TRUE)
View(tute1)
```

b. Convert the data to time series

```{r}
mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```

(The [,-1] removes the first column which contains the quarters as we don’t need them now.)

c. Construct time series plots of each of the three series

```{r}
autoplot(mytimeseries, facets=TRUE)
```

Check what happens when you don’t include facets=TRUE.

```{r}
autoplot(mytimeseries)
```

## Question 2.3

Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

a. You can read the data into R with the following script:

```{r}
if(!file.exists("retail.xlsx")){
  download.file("https://otexts.com/fpp2/extrafiles/retail.xlsx", "retail.xlsx")
}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

The second argument (skip=1) is required because the Excel sheet has two header rows.

b. Select one of the time series as follows (but replace the column name with your own chosen column):

```{r}
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
```

c. Explore your chosen retail time series using the following functions:

`autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()`

```{r}
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)
```

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

**There is a clear seasonality increase in retail sales from October to the end of the year.  This is the Christmas shopping season.  There is also a trend of increasing retail sales over time.  The trend has been rising untill the 2000's where it flatened out for roughly a decade.  Since 2010 it looks like the trend increases again.**

## Question 2.6

Use the following graphics functions: `autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

### hsales

* Can you spot any seasonality, cyclicity and trend?

```{r}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

* What do you learn about the series?

### usdeaths

* Can you spot any seasonality, cyclicity and trend?

```{r}
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

* What do you learn about the series?

### bricksq

* Can you spot any seasonality, cyclicity and trend?

```{r}
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)
```

* What do you learn about the series?

### sunspotarea

* Can you spot any seasonality, cyclicity and trend?

```{r}
autoplot(sunspotarea)
#ggseasonplot(sunspotarea)
#ggsubseriesplot(sunspotarea)
gglagplot(sunspotarea)
ggAcf(sunspotarea)
```

* What do you learn about the series?

### gasoline

* Can you spot any seasonality, cyclicity and trend?

```{r}
autoplot(gasoline)
ggseasonplot(gasoline)
#ggsubseriesplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline)
```

* What do you learn about the series?
