---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
library(kableExtra)
library(gridExtra)
library(readxl)
options(scipen=999)
```

## Question 8.1

Figure 8.31 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.

a. Explain the differences among these figures. Do they all indicate that the data are white noise?

<img src ="https://otexts.com/fpp2/fpp_files/figure-html/wnacfplus-1.png" alt="Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.">

Figure 8.31: Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.

b. Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

## Question 8.2

A classic example of a non-stationary series is the daily closing IBM stock price series (data set `ibmclose`). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

## Question 8.3

For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.

a. `usnetelec`

b. `usgdp`

c. `mcopper`

d. `enplanements`

e. `visitors`

## Question 8.5

For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

## Question 8.6

Use R to simulate and plot some data from simple ARIMA models.

a. Use the following R code to generate data from an AR(1) model with $ϕ_1 = 0.6$ and $σ_2 = 1$.  The process starts with $y_1=0$.

```{r}
y <- ts(numeric(100))
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
```

b. Produce a time plot for the series. How does the plot change as you change $ϕ_1$?

c. Write your own code to generate data from an MA(1) model with $θ_1=0.6$ and $σ_2=1$.

d. Produce a time plot for the series. How does the plot change as you change $θ_1$?  

e. Generate data from an ARMA(1,1) model with $ϕ_1=0.6$, $θ_1=0.6$ and $σ_2=1$.

f. Generate data from an AR(2) model with $ϕ_1=−0.8$, $ϕ_2=0.3$ and $σ_2=1$. (Note that these parameters will give a non-stationary series.)

g. Graph the latter two series and compare them.

## Question 8.7

Consider `wmurders`, the number of women murdered each year (per 100,000 standard population) in the United States.

a. By studying appropriate graphs of the series in R, find an appropriate ARIMA(p,d,q) model for these data.

b. Should you include a constant in the model? Explain.

c. Write this model in terms of the backshift operator.

d. Fit the model using R and examine the residuals. Is the model satisfactory?

e. Forecast three times ahead. Check your forecasts by hand to make sure that you know how they have been calculated.

f. Create a plot of the series with forecasts and prediction intervals for the next three periods shown.

g. Does auto.arima() give the same model you have chosen? If not, which model do you think is better?